Last glacial cycle glacier erosion potential in the Alps

### 2021-01-28 articles/erosion/alps-erosion.tex
Rework discussion of deglacial erosion.

    Paradoxically, the model results do not show an increase of modelled
    glacier erosion rates with increasing ice volume.  On the opposite, periods
    of ice advance and maximum glaciation correspond to comparatively slow
    erosion. While this result may seem paradoxical, it is in fact
    compatible with previous field-based studies that have evidence an inverse
    relationship between air temperatures and glacier erosion rates
    \citep{Koppes.etal.2015, Cook.etal.2020}.

    Deglaciation periods, on the other hand, which are characterised by warmer
    climate and above-balance glacier extent, yield peak domain-integrated
    erosion rates. Field-based studies have associated deglaciation periods to
    higher rates of glacial erosion and proglacial sedimentation, but the increase
    has often been attributed to higher meltwater availability, promoting river
    erosion, enhanced glacier sliding, and hydrofracturing (plucking) \aref. It
    is important to note, however, that such processed are not accounted for in
    the present study. In particular, deglaciation surface meltwater is assumed
    to instantly exit the glacier with no effect on glacier dynamics.

    The enhanced glacier dynamics during periods of ice retreat is here
    modelled to result from changes of air temperature alone, reported through
    diffusion and advection through the ice, thereby facilitating shear strain
    and eventually activating basal sliding. The observed feedbacks between
    surface meltwater and basal sliding would presumably enhance erosion peaks
    modelled to occur during deglaciation. Enhanced glacier erosion can be
    expected to accompany the ongoing 21st-century deglaciation phase.

>>>>

    The modelled, domain-integrated glacier erosion rates do not increase
    together with ice volume. On the opposite, periods of ice advance and
    maximum glaciation correspond to comparatively slow erosion
    (Figs.~\ref{fig:cumulative} and~\ref{fig:evolution}). While this result may
    seem paradoxical, and is modulated by the exponent, $l$, in the erosion
    power-law, it is in fact compatible with previous field-based studies that
    have evidenced an inverse relationship between air temperatures and glacier
    erosion rates \citep{Koppes.etal.2015, Cook.etal.2020}.

    Deglaciation periods, characterised by a warmer climate and above-balance
    glacier extent, yield the highest domain-integrated erosion rates of the
    glacial cycle (Figs.~\ref{fig:cumulative} and~\ref{fig:evolution}). This
    result is compatible with the field-based studies, which have evidenced
    increased glacier erosion and pro-glacial sedimentation during deglaciation
    periods \ref, including the ongoing glacier retreat \aref. Yet this
    increase has often been attributed to higher meltwater availability,
    promoting river erosion, enhanced glacier sliding, and hydrofracturing
    (plucking) \aref.

    But importantly, these hydrological processes are not included in our
    model, so that deglaciation surface meltwater is assumed to instantly exit
    the glacier with no effect on glacier dynamics. The enhanced glacier
    velocity during periods of ice retreat is here modelled to result from
    changes of air temperature alone, propagating through diffusion and
    advection into the ice, increasing the glacier's internal energy through
    temperature and englacial water content, thereby facilitating shear strain
    and eventually activating basal sliding \citep[cf.][]{Aschwanden.etal.2012,
    Seguinot.etal.2016}. Observed feedbacks between surface meltwater and basal
    sliding, not included in the model, presumably further accelerated erosion
    during deglaciation. Rapid glacier erosion can be expected to accompany the
    ongoing 21st-century deglaciation phase.
